---
title: "Homework 6"
author: "Tiana Noll-Walker"
date: "10/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")

COL <- read.csv("COL.csv")

BB2020 <- read.csv("2020bb_values.csv")


```
# Problem 1 
In your own words, describe what is measured by each of the following and give an example of a dataset that it could reasonably be applied to analyze:

### (A) Pearson Correlation
Pearson's correlation measures linear dependence between variables. It basically is measuring how close values are to lying on a straight line. 

### (B) Spearman's Rank Correlation
Spearman's correlation measures linear dependence between variables. But instead of looking at values, it measures dependence between the ranked order of variables. It basically is trying to find the position of the sorted variables. 

### (C) Contingency Table
A contingency table analyzes the relationship between pairs of catagorical variables. 
# Problem 2 
Pick and read one of the articles presented on: https://fivethirtyeight.com/tag/hollywood-taxonomy/  (each article clusters the movies starring a specific person). Provide brief responses to the following: 

### (A) Do these seem like reasonable choices to you? 
I read the Irdris Elba piece. I think that these seemed like reasonable choices.I haven't watched all those movies, but I think those four categories do a good job of summing up his career. 
### (B) What other metrics or dimensions might you use to compare movies starring the same individual?
I'd be interested to see how many times he was a leading character, and maybe compare his success with some of his co stars that are in similar roles. 
### (C) Is it clear from the text how the author selected the number of clusters? If so, do you agree with their choice? If not, does their choice seem supported by the scatterplot?  Justify your responses. 
Yes, the author explained the way they chose the four different clusters well. I agreee with their choice. I think the four different groups make a lot of sense, and I'd probably group them that way as well. The only thing I might switch around is he's had quite a few blockbusters, and I might put those in their own category. Another thing is maybe put tv acting in its own category as well. 
### (D) What positive aspects do the visualizations have? Is there anything that could be improved about the visualizations?
Well the visualizations do a good job of demonstrating and explaining the clusters. The only thing is the coloring isn't really relevant, and I don't think they put all of his movies or shows in the clusters (I could be wrong though). They also didn't label many of the movies, so it's hard to know which movies/shows they placed in what category just by looking at the graph. I do like the x axis being the rating and the y axis being the box office gross. 

# Problem 3

The basketball data we looked at in class has been loaded in as a dataframe called BB2020 in the first cell. 

### (A)  Construct a correlation table for  columns c(1,5,seq(6,18,2)) of the full dataset.
```{r cor table}


pairs(BB2020[c(1,5,seq(6,18,2))])
library("GGally")
ggpairs(BB2020, columns= c(1,5,seq(6,18,2)), title="Correlations")
```
```{r cor tab}
library("corrplot")
corr<-(BB2020[,c(1,5,seq(6,18,2))])
cor(corr)
```
### (B) Which of these columns are most strongly correlated?
Rk, AdjEM
### (C)  Which of these columns are least strongly correlated?
Luck, AdjT
### (D) Construct a new correlation table for the same columns but just the rows corresponding to teams in the PAC12. 


```{r try}
library("corrplot")
#data(BB2020)
p12<- BB2020$Conf=="P12"
cor_data<-(BB2020[p12,c(1,5,seq(6,18,2))])
ggpairs(cor_data, title="Correlations")

```
```{r table}
BB2020[BB2020$Conf=="P12",c(1,5,seq(6,18,2))]
cor(cor_data)
```

### (E) Which of these columns are most strongly correlated in the new table?
Rk, AdjEm

### (F) Which of these columns are least strongly correlated in the new table?
Rk and SOSOppO
### (G) What differences do you notice between the two correlation matrices?
In the first matrix, 
# Problem 4
The cost of living data we looked at in class has been loaded in as a dataframe called COL in the first cell. 

### (A) Make a scatter plot matrix of the numeric columns
```{r string}
str(COL)
COL_numeric = select(COL,-"City")
str(COL_numeric)
pairs(COL_numeric)
```
```{r lower}
pairs(COL_numeric,lower.panel = NULL)
```
### (B) Write a brief (no more than three sentences) summary of what you observe in the plot in (A)
There are definitely a few big outliers, especially corresponding to the average rent and average disposable income. There's at least one big outlier in every category. Cinema is very linear and tightly clustered, and the rest of the categories are not very clustered and they don't have super strong correlations. 

### (C) Choose a single subplot that seems most interesting to you and make a separate scatterplot of just those two columns with the points colored by the income value. 

```{r sub}
library(GGally)
library(ggplot2)
pairs(~Avg.Rent+Cinema,data=COL_numeric)
```

+
```

```
### (D) Write a brief (no more than two sentences) summary of what you observed in the plot in (C)
The correlation between average rent and cinema is pretty weak. Also, there is one large outlier in cinema, and it is not very linear. 
